---
title: "p8105_hw1_mm4963"
author: "Melanie Mayer"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Question 1

Below is the code to create the data frame for Question 1. 

```{r q1_df}
q1_df = tibble(
  vec_numeric = runif(10, 0, 5),
  vec_logical = vec_numeric > 2,
  vec_char = c("hello", "i", "do", "not", "know", "what", "to", "write", "here", "soooo"),
  vec_factor = factor(c("level1", "level1", "level1", "level1", "level1", "level2", "level2",
                        "level2", "level2", "level2"))
)
  

```

Below I will try to get the mean for each variable in the data frame created.

```{r q1_means, warning = FALSE}

mean(q1_df$vec_numeric)
mean(q1_df$vec_logical)
mean(q1_df$vec_char)
mean(q1_df$vec_factor)

```

The numeric vector allows you to take the mean because it holds numeric values. The logical vector also allows you to take the mean because TRUE holds an underlying 1 value and FALSE holds an underlying 0 value and therefore the mean of the vector relies on these values. The character vector and the factor vector do not hold any numerical values and are therefore unable to produce a mean, resulting in an NA value. 

